### 2024-09-07 cognitive-bias.qmd
20240907 - base rate example
20240907 - updates

The *representativess heuristic* refers to the tendency for a person's judgments or predictions about individuals to be made based on how similar the individual is to (i.e., how closely the individual resembles) the person's existing mental prototypes.
In this way, the representativeness heuristic involves stereotyping, which is not a statistically optimal approach to predictions [@Kahneman2011].
Moreover, the representativness heuristic does not adequately consider the quality of the evidence [@Kahneman2011].
It is thus important to question the [predictive accuracy](#sec-predictiveValidity) of the evidence for the particular prediction question.

@@

The confidence that people tend to have in their predictions depends less and the quantity and quality of the evidence for their predictions and more on the quality of the narrative they can tell about the (little) information they have immediate access to [@Kahneman2011].
For instance, people tend not to consider other relevant, important factors that they do not have immediate access to, consistent with the [WYSIATI heuristic](#sec-heuristicsWYSIATI).

>>>>

The *representativess heuristic* refers to the tendency for a person's judgments or predictions about individuals to be made based on how similar the individual is to (i.e., how closely the individual resembles) the person's existing mental prototypes.
In this way, the representativeness heuristic involves stereotyping, which is not a statistically optimal approach to predictions [@Kahneman2011].
Stereotypes are statements about groups or categories that are at least tentatively accepted as facts about every member of that group or category [@Kahneman2011].
That is, a (perhaps common) characteristic of a group is attributed to each member of that group.
In addition, the representativness heuristic does not adequately consider the quality of the evidence [@Kahneman2011].
It is thus important to question the [predictive accuracy](#sec-predictiveValidity) of the evidence for the particular prediction question.

@@

The confidence that people tend to have in their predictions depends less and the quantity and quality of the evidence for their predictions and more on the quality of the narrative they can tell about the (little) information they have immediate access to [@Kahneman2011].
For instance, people tend not to consider other relevant, important factors that they do not have immediate access to, consistent with the [WYSIATI heuristic](#sec-heuristicsWYSIATI).
Instead, people tend to focus on the cohesiveness and plausibility of the story, which differs from its probability [@Kahneman2011].
